---
title: "Exploring data-set: Categorical Analysis"
author: "Olivier PHILIPPE"
date: "26 June 2015"
output: html_document
---

```{r echo=FALSE, results='hide'}
	# Setting up the directory
	(WD <- getwd())
	if (!is.null(WD)) setwd(WD)
```

```{r echo=FALSE}
	# Read chunk from external file and do not launch the code
	read_chunk('data_generation.R')
```

```{r prepare_df, echo=FALSE}

	# prepare the dataset
```

```{r recode_last_var, echo=FALSE}
	# recode last variables
```


# 1. Introduction
This survey aims to give a broad picture of research software use and practice within the Research communities in United-Kingdom. 

```{r Q3bUni, fig.width=20, fig.height=10, echo=FALSE}
```

## 1.1. Data Collection
The survey contacted 1,000 randomly selected researchers at each of 15 Russell Group universities by email. From the 15,000 invitations to complete the survey, we received 417 responses – a rate of 3% which is fairly normal for a blind survey. 

We asked people specifically about “research software” which we defined as:

    Software that is used to generate, process or analyse results that you intend to appear in a publication (either in a journal, conference paper, monograph, book or thesis). Research software can be anything from a few lines of code written by yourself, to a professionally developed software package. Software that does not generate, process or analyse results - such as word processing software, or the use of a web search - does not count as ‘reseakhrch software’ for the purposes of this survey.

The responses were collected with Google Forms. 
The data collected during this survey is available for download from Zenodo [INSERT URL], as well as the entire code source of the data cleansing and data analysis [INSERT URL]. It is licensed under a Creative Commons by Attribution licence (attribution to The University of Edinburgh on behalf of the Software Sustainability Institute).

```{r E2vsQ8, echo=FALSE}
```
